---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  cache = TRUE,
  collapse = TRUE,
  comment = "#>"
)

# Required packages
library(ggplot2)

# Function to generate three concentric circles over a grid of points
genCircles <- function(n = 500L, radii = c(1, 2, 5)) {
  angles <- seq(from = 0, to = 2 * pi, length = n)
  names(radii) <- 1:3
  plyr::ldply(radii, .fun = function(r) {
    data.frame(x = r * cos(angles),
               y = r * sin(angles))
  }, .id = "circle")
}

circles <- genCircles()
circle1 <- circles[circles$circle == 1, ]
circle2 <- circles[circles$circle == 2, ]
circle3 <- circles[circles$circle == 3, ]

# Function to simulate data frame of random triangles
simTri <- function(n = 1L, radii = c(1, 2, 3)) {
  plyr::rdply(n, {
    angles <- runif(3, min = 0, max = 2 * pi)
    data.frame(x = radii * cos(angles),
               y = radii * sin(angles))
  }, .id = "triangle")
}

# Function to determine if a triangle is obtuse
isObtuse <- function(triangle) {
  
  x <- triangle$x
  y <- triangle$y
  
  # Lengths
  aa <- sqrt((x[1]-x[2])^2 + (y[1]-y[2])^2)
  bb <- sqrt((x[1]-x[3])^2 + (y[1]-y[3])^2)
  cc <- sqrt((x[2]-x[3])^2 + (y[2]-y[3])^2)
  
  # Test if each triangle is obtuse
  if (aa^2 + bb^2 < cc^2 | aa^2 + cc^2 < bb^2 | bb^2 + cc^2 < aa^2) {
    TRUE
  } else {
    FALSE
  }
}
```

# concentric

What is the probability that a _random triangle_ is obtuse? The answer, depends on how you define a random trinagle. In the most general case, the answer is $1/4$. To obtain this answer, one can use the fact that, for any triangle $T$, there exists some circle $C$ such that the vertices of $T$ all lie on its circumference. Hence, the problem is equivalent to finding the probability that the points randomly distirbuted on the circumference of a circle form an obtuse triangle.

A more challenging problem, inspired by the question above, is the following. Concider three concenctric circles $C_1$, $C_2$, and $C_3$ with radii $r_1$, $r_2$, and $r_3$, respectively, and that $r_1 < r_2 < r3$. What is the probability that three randomly selected points, one on each circle, forms an obtuse triagle? In other words, what is the probability that a random triangle is obtuse if each vertex must lie on the circumference of a different concentric circle? Some simulated examples (with $r_1 = 1$, $r_2 = 2$, and $r_3 = 5$) are illustrated in the figure below.

```{r examples, echo=FALSE, fig.width=6, fig.asp=1, out.width="70%", fig.align="center"}
# Simulate some random triangles
set.seed(101) 
triangles <- simTri(n = 16, radii = c(1, 2, 5))

# Plot random triangles
ggplot(triangles, aes(x = x, y = y, fill = as.factor(triangle))) + 
  geom_polygon() + 
  theme_light() +
  facet_wrap( ~ triangle, nrow = 4) +
  annotate("path", x = circle1$x, y = circle1$y) +
  annotate("path", x = circle2$x, y = circle2$y) +
  annotate("path", x = circle3$x, y = circle3$y) +
  geom_hline(aes(yintercept = 0), alpha = 0.25) +
  geom_vline(aes(xintercept = 0), alpha = 0.25) +
  theme(legend.position = "none",
        strip.background = element_blank(), 
        strip.text = element_blank()) +
  xlab("") +
  ylab("")
```

```{r proportion}
# Simulate some random triangles
set.seed(101) 
triangles <- simTri(n = 1000L, radii = c(1, 1, 5))

# What proportion are obtuse?
mean(plyr::daply(triangles, .variables = "triangle", isObtuse))
(4*5 - 1) / (4 * 5)
```

```{r proportion-plot}
props <- sapply(1:10, FUN = function(x) {
  triangles <- simTri(n = 1000, radii = c(1, 1, x))
  mean(plyr::daply(triangles, .variables = "triangle", isObtuse))
})
plot(1:10, props, type = "b", pch = 19,
     xlab = expression(r[3]), ylab = "Probability")
```